Uniform sampling without any cells to count trees

### 2024-06-28 No_tree_no_grid.Rmd
TGIF

```{r}
install.packages("pracma") # double integral
library(readxl) # to read the initial file
library(data.table) # for shift
library(pracma)
library(magrittr) # for pipes
library(tibble) # for add_column
```

@@

```{r}
## First simple function: 1 given to each point
random_points$rho1 <- rep(1, n)
# Formula given in Cordy (1993), page 3
```

@@

```{r}
# Value to estimate: the area of the territory
area <- 1000*1000
# integral2(fun1, 0, 1000, 0, 1000)$Q # the same
```

```{r}
# Estimated value
est_area <- sum((random_points$rho1)/pi) # Equal to 1e+06 whatever the chosen points
```

## Second function: (x,y) $\mapsto$ (x/y) with or without normalization

```{r}
# x/y given to each point
f_ratio <- function(x,y){
  x/y
}
```

```{r}
# Normalization of the ratio
f_norm <- function(x,y){
  (x/y)/(integral2(f_ratio, 0, 1000, 0, 1000)$Q)
}
```

```{r}
# random_points$rho2 <- f_norm(random_points$x,random_points$y)
```

```{r}
## Value to estimate
# ratio_norm <- integral2(f_norm, 0, 1000, 0, 1000)$Q # = 1
```

```{r}
## Estimated value
# est_ratio_norm <- sum((random_points$rho2)/pi)
## NOT SATISFACTORY: seems too low (between 0 and 1)
## 0.5449481 e.g.
```

```{r}
## Value to estimate
# ratio_norm <- integral2(f_ratio, 0, 1000, 0, 1000)$Q # = 1
```

```{r}
# Calculation of the ratio for each point
random_points$rho2 <- f_ratio(random_points$x,random_points$y)
```

```{r}
# Value to estimate
ratio <- integral2(f_ratio, 0, 1000, 0, 1000)$Q # = 3715802
ratio
```

```{r}
# Estimated value
est_ratio <- sum((random_points$rho2)/pi) # = 2024919
est_ratio
# Seems not very satisfactory
```

## Third function: (x,y) $\mapsto$ as.numeric(x>y)

```{r}
# (x>y) given to each point
f_ind <- function(x,y){
  as.numeric(x > y)
}
```

```{r}
# Value to estimate
half_area <- integral2(f_ind, 0, 1000, 0, 1000)$Q
half_area
# 509148 if x >= y; 487316.7 if x > y
```

```{r}
random_points$rho3 <- f_ind(random_points$x,random_points$y)
```

```{r}
# Estimated value
est_ind <- sum((random_points$rho3)/pi) # = 470000
est_ind
# Seems credible
```

@@

```{r}
n <- 100 # Sample size
pi <- sum(rep((1/1000)*(1/1000), n))
```

@@

```{r}
# New variable: rho1^2
random_points$carre <- (random_points$rho1)^2
est_carre <- sum((random_points$carre)/pi)
# est_carre = est_area # Logical since rho1 = carre = 1
```

```{r}
var_area <- est_carre/pi - (1/n)*(est_area^2)
var_area
# The empirical variance is null as expected
```

```{r}
# fun1(x,y)^2 = fun1(x,y)
varTH_area <- integral2(fun1, 0, 1000, 0, 1000)$Q/pi - (1/n)*(integral2(fun1, 0, 1000, 0, 1000)$Q^2)
varTH_area
# The theoretical variance is null as expected
# Since the total is equal to the area
```

## Case of the ratio

```{r}
# New variable: rho2^2
random_points$carre2 <- (random_points$rho2)^2
est_carre2 <- sum((random_points$carre2)/pi)
# The new total is: 16571726
est_carre2

var_ratio <- est_carre2/pi - (1/n)*(est_ratio^2)
# Huge empirical variance: 124714284077
var_ratio
```

```{r}
f_ratio2 <- function(x,y){
  (x/y)^2
}
varTH_ratio <- integral2(f_ratio2, 1e-100, 1000, 1e-100, 1000)$Q/pi - (1/n)*(integral2(f_ratio, 1e-100, 1000, 1e-100, 1000)$Q^2)
varTH_ratio
# Huge theoretical variance: 3.216476e+97
# integral2(f_ratio2, 1, 2, 1, 2) # works
# integral2(f_ratio2, 0, 1000, 0, 1000) # doesn't work
```

@@

```{r}
var_ind <- est_ind/pi - (1/n)*(est_ind^2)
var_ind
# Empirical variance: 2.491e+09 (huge)
```

```{r}
# f_ind(x,y)^2 = f_ind(x,y)
varTH_ind <- integral2(f_ind, 0, 1000, 0, 1000)$Q/pi - (1/n)*(integral2(f_ind, 0, 1000, 0, 1000)$Q^2)
varTH_ind
# Theoretical variance: 2498391345, i.e. around 2.498e+09
# Close to the empirical one: great
```

@@

## First MC application step by step with the normalized ratio

@@

```{r}
for (c in 1:B) {
  col_ <- paste("est", c, sep = "")
  # With normalized ratio
  a <- f_norm(series_MC[(2*c)-1],series_MC[2*c])
  rho_MC <- rho_MC %>% add_column(col_ = a)
  col_names <- append(col_names, col_)
}
rho_MC$x.1 <- NULL
rho_MC$y.1 <- NULL
colnames(rho_MC) <- col_names
```

@@

```{r}
for (c in 1:B) {
  est_tot <- sum((rho_MC[c])/pi)
  est_MC <- append(est_MC, est_tot)
}
```

```{r}
# Empirical mean based on MC
mean(est_MC) # should be close to 1
# 1.740544
```

@@

true_vs_MC <- function(n, B, f){
  # n: sample size; B: number of iterations for MC; f: function to consider (with 2 variables, x and y)
  pi <- sum(rep((1/1000)*(1/1000), n))

@@

  # 3rd step: List of estimated totals thus obtained from each sample
  est_MC <- c()
  for (c in 1:B) {
    est_tot <- sum((rho_MC[c])/pi)
    est_MC <- append(est_MC, est_tot)
  }

  true_value <- integral2(f, 0, 1000, 0, 1000)$Q # Theoretical value
  est_value <- mean(est_MC) # Estimated value
  return(c(true_value, est_value))
}
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = fun1)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# 1e+06 1e+06 (Equality = great)
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = f_norm)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# 1.000000 2.265018
# 1.000000 2.127999
# 1.000000 4.984275
# 1.000000 6.072692
# 1.000000 1.982327
# Satisfactory?
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = f_ratio)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# 3715802 6631207
# 3715802 10413305
# 3715802 10325197
# Satisfactory?
```

```{r}
res <- true_vs_MC(n = 10000, B = 10000, f = f_norm)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# 1.000000 3.353247
# Not necessarily better with a larger sample and more replications for MC
```

>>>>

```{r}
library(readxl) # to read the initial file
library(data.table) # for shift
library(magrittr) # for pipes
library(tibble) # for add_column
```

@@

@@

```{r}
# 1 given to each sampled point
random_points$rho1 <- rep(1, n)
# Formula given in Cordy (1993), page 3
```

```{r}
# Value to estimate: the area of the territory
area <- 1000*1000
```

```{r}
# Estimated value
est_area <- sum((random_points$rho1)/p) 
# Equal to 1e+06 whatever the chosen points
```

## Second function: (x,y) $\mapsto$ as.numeric(x>y)

```{r}
## Indicator of (x>y)
f_ind <- function(x,y){
  as.numeric(x > y)
}
```

```{r}
# Value to estimate
half_area <- 1000*1000/2
```

```{r}
# 1 given to each sampled point that satisfies x>y
random_points$rho2 <- f_ind(random_points$x,random_points$y)
```

```{r}
# Estimated value
est_ind <- sum((random_points$rho2)/p)
est_ind
# 510000 e.g.
```

## Third function: (x,y) $\mapsto$ x

```{r}
# x given to each point
f_x <- function(x,y){
  x
}
```

```{r}
# Value to estimate
val_x <- 500000000
```

```{r}
# Estimated value
est_x <- sum((random_points$x)/p) # = 
est_x
# 499690876 e.g.
```

@@

@@

```{r}
# New variable: rho1^2
random_points$carre <- (random_points$rho1)^2
est_carre <- sum((random_points$carre)/p)
# est_carre = est_area # Logical since rho1 = carre = 1
```

```{r}
var_area <- est_carre/p - (1/n)*(est_area^2)
var_area
# The empirical variance is null as expected
```

```{r}
# fun1(x,y)^2 = fun1(x,y)
varTH_area <- area/p - (1/n)*(area^2)
varTH_area
# The theoretical variance is null as expected
# Since the total is equal to the area
```

@@

```{r}
var_ind <- est_ind/p - (1/n)*(est_ind^2)
var_ind
# Empirical variance: 2.491e+09 (huge)
```

```{r}
# f_ind(x,y)^2 = f_ind(x,y)
varTH_ind <- half_area/p - (1/n)*(half_area^2)
varTH_ind
# Theoretical variance: 2.5e+09, i.e. around 2.498e+09
# Close to the empirical one: great
```

## Case of the function that keeps x

```{r}
f_x2 <- function(x,y){
  x^2
}
```

```{r}
# Estimated value
est_x2 <- sum(((random_points$x)^2)/p)
est_x2
# 499690876 e.g.
```

```{r}
var_x <- est_x2/p - (1/n)*(est_x^2)
var_x
# Empirical variance: 8.407871e+14 ?!
```

```{r}
# f_ind(x,y)^2 = f_ind(x,y)
# varTH_x <- integral2(f_x, 0, 1000, 0, 1000)$Q/p - (1/n)*(integral2(f_x, 0, 1000, 0, 1000)$Q^2)
# varTH_x
# Theoretical variance: 2498391345, i.e. around 2.498e+09
# Close to the empirical one: great
```

@@

## First MC application step by step with the function that keeps x

@@

```{r}
for (c in 1:B) {
  col_ <- paste("est", c, sep = "")
  # With normalized ratio
  a <- f_x(series_MC[(2*c)-1],series_MC[2*c])
  rho_MC <- rho_MC %>% add_column(col_ = a)
  col_names <- append(col_names, col_)
}
rho_MC$x.1 <- NULL
rho_MC$y.1 <- NULL
colnames(rho_MC) <- col_names
```

@@

```{r}
for (c in 1:B) {
  est_tot <- sum((rho_MC[c])/p)
  est_MC <- append(est_MC, est_tot)
}
```

```{r}
# Empirical mean based on MC
mean(est_MC) # should be close to 5e+08
# 499121662 e.g.
```

@@

true_vs_MC <- function(n, B, f, true_value){
  # n: sample size; B: number of iterations for MC; f: function to consider (with 2 variables, x and y)
  p <- sum(rep((1/1000)*(1/1000), n))

@@

  # 3rd step: List of estimated totals thus obtained from each sample
  est_MC <- c()
  for (c in 1:B) {
    est_tot <- sum((rho_MC[c])/p)
    est_MC <- append(est_MC, est_tot)
  }

  est_value <- mean(est_MC) # Estimated value
  return(c(true_value, est_value))
}
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = fun1, true_value = 1e+06)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# 1e+06 1e+06 (Equality = great)
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = f_ind, true_value = 500000)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])
# True integral: 5e+05
# Mean of all the estimations: 499600
```

```{r}
res <- true_vs_MC(n = 1000, B = 1000, f = f_x, true_value = 5e+08)
cat("True integral:", res[1], "\nMean of all the estimations:", res[2])

#True integral: 5e+08
# Mean of all the estimations: 500189987
```

### 2024-06-28 Trees_no_grid.Rmd
TGIF

```{r}
# Radius
R <- 5
area_circle <- pi*(R^2)
```

@@

```{r}
# Uniform sampling over the whole territory except the edge
random_points$x <- runif(n, 0+R, 1000-R)
random_points$y <- runif(n, 0+R, 1000-R)
```

@@

```{r}
rm(a,b,val,col_names)
```

```{r}
# Creating a local variable (pages 10-11 in Trinh et alii (2024))
local_var <- c()
for (placette in 1:n) {
  local_var <- append(local_var, sum(trees_in[trees_in[placette] != 0,]/area_circle))
}

est_nb <- sum((local_var)/p)
cat("An estimation of the total number of trees based on the sample:", est_nb)
# 29921.13 e.g.
```

@@

  # Intermediary step: which trees are inside the circle?
  trees_in <- data.frame(useless = rep(0,30942))
  val <- is_point_in_circle(x1, x2, trees$x, trees$y, R)
  trees_in <- trees_in %>% add_column(new_col = val)
  trees_in[1] <- NULL

  # Creation of the local variable
  local_var <- sum(trees_in[new_col != 0,]/area_circle)
  return(local_var) # A single number
}
```

@@

  # 1st step: generating n samples
  series_MC <- data.frame(x.1 = runif(n, 0+R, 1000-R), y.1 = runif(n, 0+R, 1000-R))
  col_names <- c("x1", "y1")
  for (b in 2:B) {
    x <- paste("x", b, sep = "")
    y <- paste("y", b, sep = "")
    # Uniform sampling over the whole territory except the edge
    a1 <- runif(n, 0+R, 1000-R)
    a2 <- runif(n, 0+R, 1000-R)
    series_MC <- series_MC %>% add_column(x = a1, y = a2)
    col_names <- append(col_names, c(x, y))
  }
  colnames(series_MC) <- col_names

@@

```{r}
res <- true_vs_MC(n = 100, B = 1000, R = 5)
cat("True number of trees:", res[1], "\nMean of all the estimations (Monte Carlo):", res[2])
```

>>>>

```{r}
# Radius
R <- 5
```

@@

```{r}
# Uniform sampling over the whole territory except the edge
random_points$x <- runif(n, 0, 1000)
random_points$y <- runif(n, 0, 1000)
```

@@

```{r}
rm(a,b,val,col_names, point, new_col)
```

```{r}
# Creating a local variable (pages 10-11 in Trinh et alii (2024))
local_var <- c()
source("~/work/Sim_article2024/Inc_area.R")
for (placette in 1:n) {
  local_var <- append(local_var, sum(trees_in[trees_in[placette] != 0,]/inc_area(c(random_points[placette,]$x, random_points[placette,]$y),0,0,1000,1000,R)))
}

est_nb <- sum((local_var)/p)
cat("An estimation of the total number of trees based on the sample:", est_nb)
# 26228.73 e.g.
# 38197.19 e.g.
```

```{r}
rm(placette)
```

@@

  # Intermediary step: which trees are inside the circle?
  trees_in <- data.frame(useless = rep(0,30942))
  new_col <- "in_or_out"
  val <- is_point_in_circle(x1, x2, trees$x, trees$y, R)
  trees_in <- trees_in %>% add_column(new_col = val)
  trees_in[1] <- NULL

  # Creation of the local variable
  local_var <- sum(trees_in[new_col != 0,]/inc_area(c(x1, x2),0,0,1000,1000,R))
  return(local_var) # A single number
}
```

@@

  # 1st step: generating n samples
  series_MC <- data.frame(x.1 = runif(n, 0, 1000), y.1 = runif(n, 0, 1000))
  col_names <- c("x1", "y1")
  for (b in 2:B) {
    x <- paste("x", b, sep = "")
    y <- paste("y", b, sep = "")
    # Uniform sampling over the whole territory except the edge
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    series_MC <- series_MC %>% add_column(x = a1, y = a2)
    col_names <- append(col_names, c(x, y))
  }
  colnames(series_MC) <- col_names

@@

```{r}
res <- true_vs_MC(n = 100, B = 1000, R = 5)
cat("True number of trees:", res[1], "\nMean of all the estimations (Monte Carlo):", res[2])
# True number of trees: 30942 
# Mean of all the estimations (Monte Carlo): 30954.24
```
